---
title: "Task 3: Bootstrapped Confidence Interval for Proportions"
author: "Elmera Azadpour"
date: "3/3/2021"
output: 
  html_document: 
    theme: paper
    toc: true
    toc_float: true
    code_folding: hide
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(boot)
library(patchwork)
```

### Load vector of data 
```{r}
nb <- rnorm(mean=.61, sd=1, n=36)
plot(nb)
nb = data.frame(nb)
mean(nb$nb)

a = replicate(22, 1)
b = replicate(14, 0)
c = c(a,b)
mean(c)
nb <- as.data.frame(c) %>%  
  rename("n" = "c") 

nb$n = as.numeric(nb$n)

```

## Summary statistics from samples:
```{r}
mean(nb$n)
t.test(nb$n)
hist(nb$n)
```

### Creating bootstrapping 
```{r}
# first, create a function that will calculate the mean of each bootstrapped sample
mean_fun <- function (x,i) {mean(x[i])}

# then, get the vector of salinity (salinity$sal)
nonbinary_nc <- nb$n

# create 100 bootstrap samples by resampling from the salinity vector, using the function you created to calculate the mean of each
nonbinaryboot_100 <- boot(nonbinary_nc, 
                    statistic = mean_fun,
                    R = 100)

# for comparison, let's also create 10000 bootstrap samples with replacement
nonbinaryboot_10k <- boot(nonbinary_nc, 
                 statistic = mean_fun, 
                 R = 10000)

# for comparison, let's also create 100,000 bootstrap samples with replacement
nonbinaryboot_100000 <- boot(nonbinary_nc, 
                 statistic = mean_fun, 
                 R = 100000)

# check out output of bootstrap:
nonbinaryboot_100
nonbinaryboot_10k
nonbinaryboot_100000

# in order to get the same output we would need to use set.seed() 
```

### Look at bootstrap samples using `t0` (to see oringal sample mean) on `t` (to see the means of bootstraps)
```{r, warning =FALSE, results="hide"}
nonbinaryboot_100$t0 # the original sample mean
nonbinaryboot_10k$t # the means for each of the 100 bootstrap samples
nonbinaryboot_100000$t # more means

# make vectors of bootstrap sample means a data frame (so ggplot will deal with it).
nonbinaryboot_100_df <- data.frame(bs_mean = nonbinaryboot_100$t)
nonbinaryboot_10k_df <- data.frame(bs_mean = nonbinaryboot_10k$t)
nonbinaryboot_100000_df <- data.frame(bs_mean = nonbinaryboot_100000$t)

## ggplot the bootstrap sample mediansL
# the histogram of the original sample:
p1 <- ggplot(data = nb, aes(x = n)) +
  geom_histogram()

# histogram of 100 bootstrap sample means:
p2 <- ggplot(data = nonbinaryboot_100_df, aes(x = bs_mean)) +
  geom_histogram()

# histogram of 10k bootstrap sample means:
p3 <- ggplot(data = nonbinaryboot_10k_df, aes(x = bs_mean)) +
  geom_histogram()

p4 <- ggplot(data = nonbinaryboot_100000_df, aes(x = bs_mean)) +
  geom_histogram()

# use {patchwork} 
(p1 + p2) / (p3 + p4) & theme_minimal()
```

### Use `boot.ci()` to find the CI for the bootstrapped distribution (here, with the 10k & 100,000 bootstrapped means):
```{r}
boot.ci(nonbinaryboot_10k, conf = 0.95, type = c("perc"))
boot.ci(nonbinaryboot_100000, conf = 0.95, type = c("perc"))
```


### Bootstrapped CI:
+ Based on the `boots.ci` output, we find that the CI percentile ranges from 0.44 to 0.78 or 44 to 78%. For a 95% confidence interval, that means we expect that 95 of 100 calculated confidence intervals will contain the actual population parameter. 