---
title: 'Task 1: Agglomerative hierarchical clustering'
author: "Elmera Azadpour"
date: "3/2/2021"
output: 
  html_document: 
    theme: paper
    toc: true
    toc_float: true
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
library(janitor)
# Packages for cluster analysis:
library(NbClust)
library(cluster)
library(factoextra)
library(dendextend)
library(ggdendro)
```

### Read in data
```{r, results = "hide"}
stream_chem <- read_csv(here("data","sbc_lter_registered_stream_chemistry.csv")) %>% 
  clean_names() %>% 
  na_if(-999.0) %>% 
  group_by(site_code) %>% 
  summarise(across(nh4_u_m:spec_cond_u_spercm, mean, na.rm = TRUE)) %>% 
  drop_na()
```

### Scale the data 
```{r}
## scale the numeric variables
stream_chem_scaled <- stream_chem %>% 
  select(2:11) %>% 
  scale()

## make rownames the site code names from original data set
rownames(stream_chem_scaled) <- stream_chem$site_code

# Compute dissimilarity values (Euclidean distances):
stream_euc_distance <- dist(stream_chem_scaled, method = "euclidean")

# Check out the output:
# View(euc_distance)
```

### Perform hierarchical clustering by complete linkage with `stats::hclust()`
```{r}
# Hierarchical clustering (complete linkage)
stream_hc_complete <- hclust(stream_euc_distance, method = "complete" )

# Plot it (base plot):
plot(stream_hc_complete, cex = 0.6, hang = -1)
```

### Convert to class: dendogram
```{r}
stream_dend_complete <- as.dendrogram(stream_hc_complete)
```

### Plot using `ggdendrogram()`, a `ggplot` wrapper: 
```{r}
# ggplot 
ggdendrogram(stream_dend_complete, 
             rotate = TRUE) +
  theme_minimal() +
  labs(x = "Santa Barbara area watershed site code") +
  ylab(NULL)
```


### Citation/Link to Data: 
+ Link to the data: https://drive.google.com/file/d/16rYLBi-CgvjcWhOsX1SLdD9HHUMP9m2l/view?usp=sharing

+ Data & Metadata source: 
SBC LTER: Stream chemistry in the Santa Barbara Coastal drainage area, ongoing since 2000
Creators: Santa Barbara Coastal LTER, & Melack, John M
Citation: Santa Barbara Coastal LTER and J. Melack. 2019. SBC LTER: Land: Stream chemistry in the Santa Barbara Coastal drainage area, ongoing since 2000 ver 16. Environmental Data Initiative. https://doi.org/10.6073/pasta/67a558a24ceed9a0a5bf5e46ab841174. 